---
title: "Clarabel Solver Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Clarabel Solver Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(clarabel)
```

## Introduction

These examples are from the original [Clarabel
documentation](https://github.com/oxfordcontrol/ClarabelDocs). 

## 1. Basic Quadratic Program Example

Suppose that we want to solve the following 2-dimensional quadratic programming problem:

$$
\begin{array}{ll} \text{minimize} &  3x_1^2 + 2x_2^2 - x_1 - 4x_2\\
\text{subject to} &  -1 \leq x \leq 1, ~ x_1 = 2x_2
\end{array}
$$

We will show how to solve this problem using Clarabel in R.

The first step is to put the problem data into the standard form expected by the solver.

### 1.1. Objective function

The Clarabel solver's default configuration expects problem data in the form $\frac{1}{2}x^\top P x + q^\top x$.   
We therefore define the objective function data as

$$
P = 2 \cdot \begin{bmatrix} 3 & 0 \\ 0 & 2\end{bmatrix}
\mbox{ and }
q = \begin{bmatrix} -1 \\ -4\end{bmatrix}.
$$

### 1.2. Constraints

The solver's default configuration expects constraints in the form
$Ax + s = b$, where $s \in \mathcal{K}$ for some composite cone
$\mathcal{K}$.  We have 1 equality constraint and 4 inequalities, so
we require the first element of $s$ to be zero (i.e. the first
constraint will correspond to the equality) and all other elements
$s_i \ge 0$.  Our cone constraint on $s$ is therefore

$$
s \in \mathcal K = \{0\}^1 \times \mathbb{R}^4_{\ge 0}.
$$

Define the constraint data as

$$
A =
\begin{bmatrix} 1 & -2 \\ 1 & 0 \\ 0 & 1 \\ -1 & 0 \\ 0 & -1\end{bmatrix}
\mbox{ and }
b=\begin{bmatrix} 0 \\ 1 \\ 1 \\ 1 \\ 1 \end{bmatrix}.
$$

Note that Clarabel expects inputs in Compressed Sparse Column (CSC) format for both $P$ and $A$.

### 1.3. Solution

```{r}
P <- Matrix::Matrix(2 * c(3, 0, 0, 2), nrow = 2, ncol = 2, sparse = TRUE)
P <- as(P, "symmetricMatrix")
q <- c(-1, -4)
A <- Matrix::Matrix(c(1, 1, 0, -1, 0, -2, 0, 1, 0, -1), ncol = 2, sparse = TRUE)
b <- c(0, 1, 1, 1, 1)
cones <- c(z = 1L, l = 4L)  ## 1 equality and 4 inequalities, in order
s <- clarabel(A = A, b = b, q = q, P = P, cone = cones)
cat(sprintf("Solution (x1, x2) = (%f, %f)\n", s$x[1], s$x[2]))
```

### 2. Basic SOCP Example

We want to solve the following 2-dimensional optimization problem:

$$
\begin{array}{ll} \text{minimize} & x_2^2\\[2ex]
\text{subject to} &  \left\|\begin{pmatrix} 2x_1 \\ x_2 \end{pmatrix}
- \begin{pmatrix} 2 \\ 2 \end{pmatrix}\right\|_2 \le 1
\end{array}
$$


### 2.1. Objective function

The Clarabel solver's default configuration expects problem data in the form $\frac{1}{2}x^\top P x + q^\top x$.   
We therefore define the objective function data as


$$
P = 2 \cdot \begin{bmatrix} 0 & 0 \\ 0 & 1\end{bmatrix}
\mbox{ and }
q = \begin{bmatrix} 0 \\ 0\end{bmatrix}.
$$


### 2.2. Constraints

The solver's default configuration expects constraints in the form $Ax + s = b$, where $s \in \mathcal{K}$ for some
composite cone $\mathcal{K}$.   We have a single constraint on the 2-norm of a vector, so we rewrite


$$
\left\|\begin{pmatrix} 2x_1 \\ x_2 \end{pmatrix} - \begin{pmatrix} 2 \\ 2 \end{pmatrix}\right\|_2 \le 1
\quad \Longleftrightarrow \quad
\begin{pmatrix} 1 \\ 2x_1 - 2\\ x_2 - 2 \end{pmatrix} \in \mathcal{K}_{SOC}
$$
which puts our constraint in the form $b - Ax \in \mathcal{K}_{SOC}$.

### 2.3. Solution

```{r}
P <- Matrix::Matrix(2 * c(0, 0, 0, 1), nrow = 2, ncol = 2, sparse = TRUE)
P <- as(P, "symmetricMatrix")
q <- c(0, 0)
A <- Matrix::Matrix(c(0, -2.0, 0, 0, 0, 1.0), nrow = 3, ncol = 2, sparse = TRUE)
b <- c(1, -2, -2)
cones <- c(q = 3L)
s <- clarabel(A = A, b = b, q = q, P = P, cone = cones)
cat(sprintf("Solution (x1, x2) = (%f, %f)\n", s$x[1], s$x[2]))
```
